Large-scale harmonic patterns clustering

### 2010-03-20 Paper/Paper.tex
improved

\section{Visualization}
The goal here is to give an intuition to the reader of what kind of
typical patterns do we get. We also want to have a feel for what the
clusters look like, how many boring sustaned notes are encoded in the
codebook, can the codebook describes the complexity of the music space, etc.

@@

\section{Further Experiments}\label{sec:exps2}

\subsection{Song Segmentation}
We explained that we used bars as defined by the EchoNest algorithm to
train our patterns. Let's assume once again that those bars are
reasonable. We investigate whether we could recover the bar segmentation
of new songs from our codebook. It would helps us assert that codes
encodes something related to bars, for instance a strong beat at the
beginning.

See Table \ref{tab:offset}.

@@

\subsection{Artist Recognition}
Let us be clear, the clustering method proposed in this work is not
intended to perform well at artist recognition. An obvious fact is that
we lose all timbral information which would be usefull. However, we can
imagine that some artist use more often some patterns than others.
Best accuracy reported is around $54.5\%$. Our result is $23.4\%$. Random
is around $5\%$.

See Figure \ref{fig:conf_mat}.

>>>>

\section{Visualization} \label{sec:visu}
The goal here is to give an intuition to the reader of what kind of
typical patterns do we get. We also want to have a feel for what the
clusters look like, how many boring sustaned notes are encoded in the
codebook, can the codebook describes the complexity of the music space, etc.

@@

\section{Further Experiments}\label{sec:exps2}
We have shown the basic properties of our clustering algorithm in
Section \ref{sec:experiments}. In Section \ref{sec:visu}, we showed
how to visualize the codebook and the clusters. We know move on to
show other applications that can take advantage of our clustering.
In the next two subsections, we talk about song segmentation and
artist recognition. Our algorithm was not devised for any of these tasks,
but some information contained in the centroids can be naturally used.

\subsection{Song Segmentation}
We explained that we used bars as defined by the EchoNest algorithm to
train our patterns. Let's assume once again that those bars are
reasonable. We investigate whether we could recover the bar segmentation
of new songs from our codebook. It would helps us assert that codes
encodes something related to bars, for instance a strong beat at the
beginning. Results are shown \ref{tab:offset}. We find the right segment
with an accuracy of $62\%$ where random is $25\%$. Details of the
experiment follow.

We train a codebook of size $100$ on bars resized to $4$ beats. Then,
we take the longest sequence of bars of $4$ beats in the test set.
As we mentioned before, $4$ beats is the most common (by far!) bar size
in terms of beats according to the EchoNest API. We then encode this
sequence using an offset of $0$, $1$, $2$ or $3$ beats. The best offset
is the one that gives the encoding with the lowest distortion. The
good answer is an offset of $0$.

@@

\subsection{Artist Recognition}
Let us be clear, the clustering method proposed in this work is not
intended to perform well at artist recognition. An obvious fact is that
we lose all timbral information which would be usefull. However, we can
imagine that some artist use more often some patterns than others.
Best accuracy reported is around $54.5\%$. Our result is $23.4\%$. Random
is around $5\%$. The confusion matrix can be seen in Figure \ref{fig:conf_mat}.
Details of the experiment follow.

All songs in the artist dataset are encoded as an histogram of the codes
used in that song. The frequency values are normalized by the number
of patterns in the song. We test each song in a leave-one-out setting.
Leaving aside the test song, we represent the $20$ artists by the average of
the song histograms of that artist. The test song is compared to
the histogram of each artist using euclidean distance.
